---
title: "Poker"
author: "Jeswanth Yadagani, Kavita Anant, Prutha Parmar"
date: "01/12/2019"
output:
  html_document:
    df_print: paged
---


### Load Data
```{r Load_Data}
library(ModelMetrics)
train_cols <- c('S1','C1','S2','C2','S3','C3','S4','C4','S5','C5','Class')
train_data <- read.table('poker-hand-training-true.data', sep = ",", col.names=train_cols)
test_data <- read.table('poker-hand-testing.data', sep = ",", col.names=train_cols)
X_train <- train_data[,1:10]
Y_train <- train_data[,11]
X_test <- test_data[1:25000,1:10]
k = 11
X_test <- test_data[(k*25000+1):((k+1)*25000),1:10]
Y_test <- test_data[(k*25000+1):((k+1)*25000),11]
```

```{r EDA}
plot(c(1:length(Y_test)),Y_test,type='p',col='red',main='Test dataset',xlab='Test data points',ylab='classes')
plot(c(1:length(Y_train)),Y_train,type='p',col='red',main='Train dataset',xlab='Train data points',ylab='classes')
```

### Feature Engineering
```{r Feature Engineering}

c_sort <- function(X){
  for (i in c(1:dim(X)[1])){
    X_temp = X[i,]
    c_rank = rank(c(X[i,]$C1,X[i,]$C2,X[i,]$C3,X[i,]$C4,X[i,]$C5),ties.method = 'first')
    k = 1
    for (j in c_rank){
      X[i,2*j - 1] = X_temp[2*k - 1]
      X[i,2*j] = X_temp[2*k]
      k = k+1
    }
  }
  X
}

s_sort <- function(X){
  for (i in c(1:dim(X)[1])){
    X_temp = X[i,]
    s_rank = rank(c(X[i,]$S1,X[i,]$S2,X[i,]$S3,X[i,]$S4,X[i,]$S5),ties.method = 'first')
    k = 1
    for (j in s_rank){
      X[i,2*j - 1] = X_temp[2*k - 1]
      X[i,2*j] = X_temp[2*k]
      k = k+1
    }
  }
  X
}

max_c_and_s <- function(X){
  max_s = data.frame(n_suite=NA)
  max_c = data.frame(n_class=NA)
  for (i in c(1:dim(X)[1])){
    max_s[i,1] <- max(tabulate(c(X[i,]$S1,X[i,]$S2,X[i,]$S3,X[i,]$S4,X[i,]$S5), nbins = 4))
    max_c[i,1] <- max(tabulate(c(X[i,]$C1,X[i,]$C2,X[i,]$C3,X[i,]$C4,X[i,]$C5), nbins = 13))
  }
  res1 <- cbind(X,max_s)
  res <- cbind(res1,max_c)
  res
}
```

#### Sort by suite
```{r}
s_X_train <- s_sort(X_train)
s_X_test <- s_sort(X_test)
```

#### Sort by class
```{r}

c_X_train <- c_sort(X_train)
c_X_test <- c_sort(X_test)
```

#### Add columns maximum number of Classes and maximum number of Suites
```{r}
f_s_X_train <- max_c_and_s(s_X_train)
f_s_X_test <- max_c_and_s(s_X_test)
f_c_X_train <- max_c_and_s(c_X_train)
f_c_X_test <- max_c_and_s(c_X_test)

```

```{r barplot}
bar_plot <-function(y_act,y_pred,title){
  leg=c('Actual','Predicted')
  plot_table=t(data.frame(tabulate(y_act+1),tabulate(y_pred+1,nbins=10)))
  colnames(plot_table) <- c(0:9)
  barplot(data.matrix(plot_table),main=title,xlab='Rank',ylab='count',col = c('blue','red'),legend=leg,beside=TRUE)
}

```

### Linear Regression
```{r linear regression analyse}
lin_regression<-function(X_train,X_test,Y_train,Y_test,title){
lm1 <- lm(Y_train~.,data=cbind(X_train,Y_train))
pred_class_train <- round(predict(lm1,newdata = X_train))
pred_class_test <- round(predict(lm1,newdata = X_test))
table(Y_train,pred_class_train)
table(Y_test,pred_class_test)
lr_train_error = ce(Y_train,pred_class_train)
lr_test_error = ce(Y_test,pred_class_test)
#plot(c(1:length(Y_test)),Y_test,type='p',col='red')
#lines(c(1:length(pred_class_test)),pred_class_test,type='p',col='green')
bar_plot(Y_test,pred_class_test, title)
c(lr_train_error,lr_test_error)
}
```

### Tree 
```{r tree}
library(tree)
tree_reg<-function(X_train,X_test,Y_train,Y_test,title){
tree1 <- tree(Y_train~.,data=cbind(X_train,Y_train))
tree_pred_train <- round(predict(tree1,X_train))
tree_pred_test <- round(predict(tree1,X_test))
table(Y_train,tree_pred_train)
table(Y_test,tree_pred_test)
tree_train_error = ce(Y_train,tree_pred_train)
tree_test_error = ce(Y_test,tree_pred_test)
#plot(c(1:length(Y_test)),Y_test,type='p',col='red')
#lines(c(1:length(tree_pred_test)),tree_pred_test,type='p',col='green')
bar_plot(Y_test,tree_pred_test, title)
c(tree_train_error,tree_test_error)
}
```

### Random Forest
```{r random forest analysis}
library(randomForest)
randForest<-function(X_train,X_test,Y_train,Y_test,title){
rf1 <- randomForest(x=X_train,y=Y_train,xtest=X_test,ytest=Y_test,importance= TRUE)
rf_pred_train <- round(rf1$predicted)
rf_pred_test <- round(rf1$test$predicted)
pred_train_tab <- table(Y_train,rf_pred_train)
pred_test_tab <- table(Y_test,rf_pred_test)
rf_train_error = ce(Y_train,rf_pred_train)
rf_test_error = ce(Y_test,rf_pred_test)
bar_plot(Y_test,rf_pred_test, title)
barplot(rf1$importance[,1],ylim=c(0:1),main=title,las=2)
c(rf_train_error,rf_test_error)
#plot(c(1:length(Y_test)),Y_test,type='p',col='red')
#lines(c(1:length(rf_pred_test)),rf_pred_test,type='p',col='green')
}
```

### SVM
```{r SVM analysis}
library(e1071)

svm_cl<-function(X_train,X_test,Y_train,Y_test,title){
svm1 <- svm(Y_train~.,data=cbind(f_c_X_train,Y_train))
svm_pred_train <- round(predict(svm1,f_c_X_train))
svm_pred_test <- round(predict(svm1,f_c_X_test))
table(Y_train,svm_pred_train)
table(Y_test,svm_pred_test)
svm_train_error = ce(Y_train,svm_pred_train)
svm_test_error = ce(Y_test,svm_pred_test)
bar_plot(Y_test,svm_pred_test, title)
c(svm_train_error,svm_test_error)
#plot(c(1:length(Y_test)),Y_test,type='p',col='red')
#lines(c(1:length(svm_pred_test)),svm_pred_test,type='p',col='green')
}
```


```{r}
```

## Without sorting

### Linear Regression
```{r}
lr_error_o=lin_regression(X_train,X_test,Y_train,Y_test,'Linear Regression-without sorting')
```

### Decision Trees
```{r}
tree_error_o=tree_reg(X_train,X_test,Y_train,Y_test,'Decision Trees-without sorting')
```

### Random Forest
```{r}
rf_error_o=randForest(X_train,X_test,Y_train,Y_test,'Random Forest-without sorting')
```

### SVM
```{r}
svm_error_o=svm_cl(X_train,X_test,Y_train,Y_test,'SVM-without sorting')

```


## Sorted by classes

### Linear Regression
```{r}
lr_error_c=lin_regression(f_c_X_train,f_c_X_test,Y_train,Y_test,'Linear Regression-Sorted by classes')
```

### Decision Trees
```{r}
tree_error_c=tree_reg(f_c_X_train,f_c_X_test,Y_train,Y_test,'Decision Trees-Sorted by classes')
```

### Random Forest
```{r}
rf_error_c=randForest(f_c_X_train,f_c_X_test,Y_train,Y_test,'Random Forest-Sorted by classes')
```

### SVM
```{r}
svm_error_c=svm_cl(f_c_X_train,f_c_X_test,Y_train,Y_test,'SVM-Sorted by classes')

```

## Sorted by suites

### Linear Regression
```{r}
lr_error_s=lin_regression(f_s_X_train,f_s_X_test,Y_train,Y_test,'Linear Regression-Sorted by suites')
```

### Decision Trees
```{r}
tree_error_s=tree_reg(f_s_X_train,f_s_X_test,Y_train,Y_test,'Decision Trees-Sorted by suites')
```

### Random Forest
```{r}
rf_error_s=randForest(f_s_X_train,f_s_X_test,Y_train,Y_test,'Random Forest-Sorted by suites')
```

### SVM
```{r}
svm_error_s=svm_cl(f_s_X_train,f_s_X_test,Y_train,Y_test,'SVM-Sorted by suites')

```
```{r}
lr_acc_o=((1-lr_error_o)*100)
tree_acc_o=((1-tree_error_o)*100)
rf_acc_o=((1-rf_error_o)*100)
svm_acc_o=((1-svm_error_o)*100)

```

```{r}
print('Without sorting')
print(lr_acc_o)
print(tree_acc_o)
print(rf_acc_o)
print(svm_acc_o)
```
```{r}
print('Sort by class')
print((1-lr_error_c)*100)
print((1-tree_error_c)*100)
print((1-rf_error_c)*100)
print((1-svm_error_c)*100)

```

```{r}
print('Sort by suite')
print((1-lr_error_s)*100)
print((1-tree_error_s)*100)
print((1-rf_error_s)*100)
print((1-svm_error_s)*100)

```

